Gender Segregation in Occupations: Preferences or Homophily?

### 2019-07-11 paper/Paper_2019_July_v1.tex
first round cuts

\abstract{Women remain concentrated in certain occupations despite the drastic increase in women's labor force participation in the U.S. since 1960. I examine whether occupations remain segregated because workers prefer to enter occupations that already employ more of their own gender. I build a model of labor supply and demand in which firms care about the gender and wages of their employees, and workers get utility from their occupation, wage, and the number of women in their occupation. Using a Bartik instrumental variables strategy, I find that women prefer to enter into female-dominated occupations, but men show no evidence of gender preference. In a world without women's gender preference, occupations would be 50 percent less segregated and the gender wage gap 18 percent smaller. However, women's preference for working with women would have to be almost twice as strong for past segregation to affect current segregation.}

@@

%... focused lit review !!!!!
% results.... include some case studies of mechanics and nurses?
% conclusion
%occupation gender segregation is a persistent feature of the labor market that continues to contribute substantially to the gender wage gap, especially as the education gap between men and women has closed and even reversed \cite{Blau2006, Blau2013, Kahn2016a}. One distinctive stylized fact is that tipping points exist in the fraction female in an occupation, which is suggestive of worker preferences over the gender composition of occupations \cite{Pan2010}.  Many factors have been identified as contributing causes of  segregation\footnote{ See \cite{Bertrand2011} for an overview and recent work by \cite{Black2010, Wiswall2013, Bronson2013, Olivieri2014a, Baker2015, Hsieh2016, Cortes2017a} }, but preferences over gender composition itself have received less empirical treatment\footnote{\citeA{Lordan2015} find evidence consistent with a female preference for working in occupations with more women, and also some evidence for male preferences for working with men.}. Such preferences are ripe for further exploration because other widely explored mechanisms, including preferences over amenities, productivity and skill differences, and employer discrimination, are unlikely explain the observed tipping patterns. This paper uses a novel approach to distinguish labor demand and labor supply side mechanisms to find evidence of worker preference over occupation gender composition. The impact of these preferences on sorting and policy can then be assessed.
%% focus is often on wage differentials more than sorting!
%%Also compensating differentials lit? Usui2008b, Usui2009a equilibrium search
%one cause of preferences over occupation gender composition could be utility cost to entering an opposite-gender dominated occupation as a minority. This utility cost could be due to identity cost \cite{Akerlof2000}, preferences over peers \cite{Pan2010}, or unobserved amenities correlated with the gender ratio\footnote{I focus on mechanisms that operates at the occupation level, rather than a preference over the sex of individual coworkers in an establishment or firm as in \cite{Zheng2013}. Unfortunately establishment level data such as LEHD lacks occupation data to address occupation choice.}. In any case such preferences could produce a degree of persistence and dependency on initial conditions in occupation sorting, in addition to the aforementioned tipping points. These forces matter if we want to know the economic efficiency and social welfare impact of both segregation, and policies meant to influence segregation. 
%If occupation sorting is based entirely on productivities and preferences over tasks and amenities, then public campaigns encouraging men to enter pink collar professions such as nursing, and women to enter silicon valley and STEM fields are misguided and could have negative consequences. However, if disutility from being a minority gender is an important factor, then incentivizing a transition to parity in sorted occupations could increase the total surplus of worker job matches, both by eroding the utility cost of segregation on the worker side, and on the job side by making easier for workers to match to the jobs at which they are most productive. 
%%There may exist people who would love to enter opposite sex occupation for wages and amenities but are held back by disutility from being the minority gender (eg. silicon valley). In addition, to echo the rhetoric of campaigns to encourage women to enter STEM and men to nursing, these people might have occupation specific talent to offer that is important for production. 
%%\cite{Polachek1981, Dustmann2007, DeLeire2004, Desai1991, glass1992, Reed1994}
%%Currently perceived unequal occupational opportunities may be met with affirmative action, wage or amenity laws, or public opinion campaigns. The impact of such policies will depend on the underlying causes of the segregation. If identity cost or stigma is the cause of segregation, then there could be room for policy to move to a more efficient equilibrium sorting. However if men and women have different productivities and preferences over tasks, then there could be negative consequences of policies to decrease segregation.
%% In addition the impact of changing amenities, such as making the workplace more flexible, could have no impact on segregation or even amplify it.
%%The entry of women into the labor force has provided a lot of variation to help disentangle the mechanisms of sorting. As women have entered, some occupations have received a level of entry that could be predicted given past shares, while others have received outsized entry and tipped female, while others have remained entirely male dominated despite the increase in female labor force participation. Since the 1990s occupation segregation has ceased to diminish \cite{Blau2013}. In the context of my model, we have reached a segregated steady state. I argue that even without changes in any fundamental aspects of the economy, eg. amenities or production technology, occupations could become unsegregated through a small degree of either random or incentivized entry by the opposite sex, enough to erode the stigma due to the identity association of the occupation. The erosion of the stigma could then have long run consequences on the occupation choice of future generations. 
%To distinguish the impact of preferences over gender ratio of occupation, already suggested in clean reduced form context in \citeA{Pan2010}, from the other common explanations of segregation such as task productivity and amenities, I propose a model of equilibrium static lifetime occupation choice paired with a Bartik and Hausman instruments\footnote{The second stage strategy is similar to \cite{Olivieri2014a}} to obtain clean variation in the fraction female and lifetime wage profiles at the occupation level. The transferable utility model is an adaptation of the insights of \citeA{Salanie2014} to the labor market setting, where wages are taken as observed transfers to provide additional identifying power.
%The flexibility of the matching model allows me to predict the impact of nudges to occupation sex ratios, wages, and amenities on long run gender segregation. Since the gender ratio of occupations emerges endogenously from worker and firm preferences, any attempt to manipulate the gender ratio could have long run dynamic consequences. The equilibrium model is used to simulate these consequences using repeated cross sections of overlapping generations of workers to predict that some occupations will remain at a stable gender ratio while others will move male or female over time.
%%To distinguish the gender identity hypothesis from other hypotheses, including that men and women cannot productively work together, have different intrinsic preferences over occupations, or different comparative advantages, I must have a model that identifies the utility of both the worker in choosing an occupation with observed and unobserved characteristics, and the payoff to a job in that occupation of hiring a worker with given observed and unobserved characteristics. I focus on an identity mechanism that operates at the occupation level, rather than a preference over the sex of individual coworkers in an establishment. In future work establishment level data could be used to disentangle theses mechanisms.
%%States are treated as independent labor markets with different sex ratios by occupation, and some occupation attributes (SEE CPS DATA**). Within states workers match to jobs, but the choice set available to both workers and employers is unobserved. Multiple equlibria could occur, and to avoid specifying an equilibrium selection mechanism, I adopt the rank score estimator approach of Fox 2010. Parameters on both sides of the market are estimated from the necessary inequalities for pairwise stability. If I had m:m matching this would be cool.

%\section{Literature Review} \label{literature}

%A number of papers have studied worker preferences for tasks and amenities by occupation. \citeA{asadullah2008} and \citeA{Reed1994} find that men and women have similar preferences over amenities using job satisfaction surveys and job transitions respectively. On the other hand, \citeA{Goldin2011} and \citeA{Goldin2014} find higher demand from female workers for flexibility and shorter hours to be important in explaining sorting and wage differentials. I find large variation in male and female preferences for different occupations, which could include any occupation amenities that are fixed over time, such as flexibility, task requirements,\footnote{There is evidence that changes in task requirements over time is important in explaining gender gaps \cite{Black2010, atalay2017}. In a robustness check I control for the time-varying task measures of \citeA{atalay2017} and find qualitatively similar results for gender preference over fraction female.} hours, or risk.

%Several papers look for preference over the gender composition of the occupation. Using job satisfaction data from the NLSY, \citeA{Usui2008} finds that women dislike male occupations and receive a compensating differential to work in them. Similarly \citeA{Lordan2015} use job satisfaction and transition data to find that even controlling for occupation task measures and firm level fraction female, there is still evidence for women disliking male occupations.\footnote{The estimates in \citeA{Lordan2015} are robust even to controls for firm level fraction female and remain even after the inclusion of occupation task measures, suggestive of at least some preference over the gender composition at the occupation level.} Unlike this literature my paper explicitly models how such a preference factors into occupation choice, including the impact of compensating differentials on equilibrium wage levels.

%The methodological contribution of my paper builds on a long legacy of literature on theoretical matching. \citeA{Shapley1972a} showed that with transferable utility the stable matching coincides with the competitive equilibrium. \citeA{Crawford1981} and \citeA{RothSotomayor} further developed mechanisms for convergence to the competitive equilibrium, and conditions for the uniqueness of the prices, or wages, that support it. More recently \citeA{Choo2006} pioneered empirical estimation of transferable utility models with an application to the marriage market. \citeA{Salanie2014a} and \citeA{Dupuy2017} have presented the theoretical possibility of using observed transfers for identification. My paper is the first to my knowledge to take this insight to an empirical application, using wages to disentangle firm and worker preferences in the labor market.

%A possible consequence of workers having preferences over groups is ``tipping", or rapid movements in segregation patterns. The literature on tipping includes the seminal work of \citeA{Card2008}, identifying tipping points for racial segregation in neighborhoods, \citeA{Caetano2017} finding tipping and multiple equilibria in the racial composition of schools, and of course \citeA{Pan2010} using regression discontinuity to find tipping points in fraction female in occupations. I focus on disentangling the mechanisms underlying tipping, and am the first to my knowledge to examine the role of equilibrium wages or prices in such a market.

%My setting is different because of wages?

%Similar papers in terms of topic and estimation strategy include \citeA{Bayer2005}, \citeA{Olivieri2014a}, and \citeA{Hsieh2013a}. \citeA{Bayer2005} develop an estimation strategy for equilibrium in group sorting by neighborhoods in the housing market. My paper is different because I have a two-sided model, and my static equilibrium is in wages rather than the fraction female. My paper is similar to the occupation choice model of \citeA{Olivieri2014a}, but her paper is a one-sided model and without endogenous wages. She also uses current period wages as opposed to lifetime wages. \citeA{Hsieh2013a} develops a simillar estimation strategy based on the distribution of observed wages, but in the context of a general equilibrium Roy model. \citeA{Hsieh2013a} assume that men and women are equally productive, in order to identify ``frictions" that rationalize gender gaps. My model allows for both productivity and preference differences between men and women, and wage heterogeneity based on job amenities, by shutting down the individual level productivity heterogeneity that is the hallmark of a Roy model.\footnote{The identification arguments outlined below would still go through allowing for worker heterogeneity in productivity, but this would result in weak identification and likely intractable estimation.}

% if there is only productivity variation, then firms only want to hire either men or women.... need both kinds of heterogeneity, which then makes it intractable.

%I allow workers to differ in productivity by gender but not at the individual level, in order to facilitate identification.\footnote{In future work I may include education as a proxy for worker skill, but I currently treat at as part of the occupation choice since the two choices are interrelated.} However I do allow for individual tastes for occupations, and unobserved heterogeneity in job amenities. Heterogeneity in job amenities is what produces wage heterogeneity within occupation and gender. The identification arguments outlined below would also work by shutting down heterogeneity in job amenities and allowing worker heterogeneity in productivity, but in this case the likelihood function becomes more complicated.\footnote{See \citeA{Taber2011} for a decomposition of heterogeneity from various sources. Recent work by \citeA{Sorkin2015a} finds that amenity heterogeneity does play a major role in producing wage variation at the firm level.}

%My model allows for both compensating differentials, and differences in productivity or discrimination, by observed covariates, which are gender and occupation.\footnote{In future work I may include education as a proxy for worker skill, though education choice is obviously endogenous to occupation choice, which is why I treat it as part of the occupation choice in the current specification.} At the individual level, my model only allows for unobserved heterogeneity in job amenities, not worker productivity. Wage heterogeneity within occupation and gender therefore emerges due to selection over job amenities. The focus of the paper is worker preferences, and identification of worker preferences through the wage distribution is facilitated by shutting down individual level heterogeneity in productivity.\footnote{The identification arguments outlined below also work by shutting down job level heterogeneity in amenities and allowing worker heterogeneity in productivity, but in this case the likelihood function is more complicated.} See \citeA{Taber2011} for a decomposition of heterogeneity from various sources. Recent work by \citeA{Sorkin2015a} finds that amenity heterogeneity does play a major role in producing wage variation at the firm level.

% not to decompose sources of wage heterogeneity as in \citeA{Taber2011}, but rather to decompose worker preferences.

%in amenities and for tractability reasons is it much easier to focus only on job level amenities as the source of wage heterogeneity.

%one reason we might expect men and women to make different occupation choices is the differential impact of fertility on men and women. \citeA{Polachek1981} sought to explain sorting using human capital theory but later evidence has not supported his hypothesis \cite{England1982, glass1992, England2005}. In estimating a structural model, \citeA{Dustmann2007} find that fertility does impact occupation choice, but does not explain all of occupation segregation.

% \citeA{Bonin2007} ?? can't find this paper!!! somethign about risk preferences?

%There is also mixed support for the idea that men and women have different preferences over occupations.  \citeA{DeLeire2004} find some evidence for different risk preferences leading to occupational sorting. On the other hand \citeA{Desai1991} and \citeA{Reed1994} do not find differences in occupation preferences across sex using choice data. Even in a survey of  preferences, \citeA{Tolbert1998} finds no sex differences. On the other hand, \citeA{Goldin2011} and \citeA{Goldin2014} do find higher demand from female workers for flexibility and shorter hours to be important in explaining sorting and wage differentials.
%
%The employer's payoff to hiring a man or a woman in each occupation may be due to both the productivity or skill set of the type \cite{blaukahn1994, Black2010, Autor2013, Yamaguchi2013} and a taste against employing the type \cite{Becker1957, Bergmann1974, blaukahn1994}. Like many papers in the labor literature \cite{blaukahn1994}, I cannot distinguish productivity and employer discrimination. The focus of the paper is instead on decomposing the worker's utility to explain the worker's occupation choice.

%\citeA{Card2008} successfully applied this framework to racial sorting across neighborhoods,

%occupational choice and wage determination are closely intertwined due to selection effects \cite{Roy1951}, and the possibility that preferences for amenities imply compensating differentials which may differ by sex \cite{Usui2008b, Goldin2011}.

%?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS?MECHANISMS

%A credible equilibrium model of wage determination is necessary to answer my question of the impact of preferences over fraction female on occupation choice because we precisely need to distinguish to what extent sorting is driven by demand side or supply side factors.

%Exploiting the full distribution of wage data in a matching model \cite{Salanie2014} I am able to distinguish whether a given type of worker is underpaid in an occupation because they prefer to work in that occupation (compensating differential), from a scenario where they are less productive or devalued in that occupation. Estimating wages as an equilibrium outcome of worker and job utility is critical for counterfactual simulations to reflect both these mechanisms. For example if I alter the fraction female in an occupation, that will impact the utility the worker receives from entering that occupation, directly impacting the wage through compensating differential. However the compensating differential impact will be mitigated by the substitution pattern of workers across jobs and into jobs previously unfilled, which depends also on the relative productivity/valuation of workers in those occupations.

%not only affects the utility of other women considering careers in science (which is my mechanism of interest) but also changes the wage structure and payoffs of other occupations as those women are no longer available to work in other occupations.

@@

In the model, career choice is a one-to-one match between a job and a worker. Workers choose an occupation at the beginning of their lives, based on their individual tastes for occupations and the wage offer for their gender in that occupation. The job choice, including any education costs that may be associated with it, is binding for life. I assume that workers do not anticipate future changes in the market that would lead them to switch occupations. With myopic workers, lowering switching cost would only lead to faster convergence to a fixed point in the fraction female by occupation. However, relaxing myopia would pose a challenge to tractability because the endogenous attribute (fraction female) would then depend on expectations over the future behavior of all agents, complicating today's choice set considerably.

The lifetime income distributions by gender and occupation needed to clear the market are determined in equilibrium to equate demand and supply, and are made up of two components: a job-specific component, and an aggregate component that varies by occupation and gender. The job-specific component reflects the idiosyncratic disutility of taking a specific job. In equilibrium firms must make workers indifferent across individual jobs within an occupation, and therefore firms perfectly compensate workers for idiosyncratic disutility. Wage heterogeneity within occupation and gender emerges from these job-specific compensating differentials.

@@

%I am not concerned with frictions such as learning and job changing early in life. The occupation choice that I take as data is the final occupation choice. I am also not concerned with short term frictions due to education, for instance the model does not include the type of argument that there are so few men in nursing because not enough men went to nursing school. In my model the choice of an occupation includes the choice of all schooling required for that occupation, and any friction of waiting for education to be completed or barriers to education is assumed away.

%In my PSID sample workers spend 63\% of the years 25-55 in their most common occupation\footnote{occupation is defined by the broadest aggregation of occ1990 for comparability}. In the simulated SIPP data this number is 51\%.

%In PSID and SIPP what is the fraction of workers that spend more than 50\% of their working-aged life in the same occupation or non-employed? How about more than 50\% of working years? Correlation between starting occupation and modal occupation? ... count number of workers for whom starting occupation matches model occupation.

@@

The identification strategy in this paper builds off empirical applications of matching models to marriage markets (e.g. \citeA{Choo2006} and \citeA{Chiappori2015}), but differs in several key respects. Because a key model parameter is how non-wage utility varies with occupation fraction female, it is critical that I separately identify non-wage utility from utility from wages. Previous literature only identifies the sum of wage and non-wage utility.

The first step to separate identification is to assume that observed wages are the transfers between job and worker that clear the market. Although some transfers in the labor market may be non-wage, such as enhanced benefits, wages are a natural first-order approximation. The difficulty is that observed transfers are the sum of an idiosyncratic and an aggregate component, as discussed above, and for counterfactuals these must be separately identified. The observed distribution of wages is also the result of optimization over these components and other model parameters to be estimated, specifically the willingness-to-pay for workers by gender and the variance of the distribution of unobserved heterogeneity in job amenities.

In the case that heterogeneity on both sides of the market is distributed extreme value, it is possible to solve for all parameters from a few key moments as suggested by \citeA{Salanie2014a}. However in my application, such a strategy has some major drawbacks. First, logit heterogeneity does not fit the observed wage distribution. Wages are typically assumed to be lognormal since they are constrained to be greater than zero and have a long right tail. Unlike the logit case, the variance of the distribution of the maximum of lognormals depends on other model parameters to be estimated, so the variance needs to be jointly estimated unlike in \citeA{Salanie2014a}.

Second, unlike in the marriage market, what it means to be ``unmatched" is not clearly defined on both sides of the market. For workers, it is clear that being unmatched means being unemployed or out of the labor force. On the job side, however, it is less clear. The number of unmatched jobs, as a primitive, would depend on an occupation production function and capacity constraints. Rather than estimate this, I choose to use data on posted vacancies that remain unfilled for some time. The decision to post a vacancy is endogenous to market conditions, so vacancy data is an imperfect proxy for unfilled jobs. In order to limit the importance of the vacancy data in my estimation, I choose a specification that in principle does not require it, by assuming non-negative profit on the job side. However in practice this relies very heavily on the functional form assumption on the tail of the distribution of job dis-amenities. As a result in most specifications I use data on vacancies by occupation imputed from the Job Openings and Labor Turnover Survey \cite{JOLTS}.

Lastly, I estimate four variance parameters in total: variance of taste for occupation for men and women, and variance of job dis-amenities for men and women. These parameters are important because they govern the elasticity of labor supply to changes in wages or other non-wage amenities such as the gender ratio in the counterfactual, as well as the demand response to changes in equilibrium wage. Unlike \citeA{Chiappori2015}, who uses many markets to identify these parameters, I use the observed wage distribution as suggested in \citeA{Salanie2014a}. However I estimate using maximum likelihood in order to use the information from the wage distribution efficiently, and because the lognormal assumption means that the scale parameters cannot be separately estimated from all other parameters, as noted above. For all these reasons maximum likelihood is the most natural strategy for my application.

@@

%fails in the application of the model to jobs and and workers because unlike in matching men and women in marriage, in the labor market we do not observe jobs that are left unmatched. Even when data on vacancies exists, vacancies are not an accurate representation of the potential number of jobs in the economy since the choice to post a vacancy is endogenous to market conditions. However transfers are observed between jobs and workers in the form of wages, and I show how this data can be leveraged to identify and estimate a matching model in the labor market context.

@@

The fact that jobs care only about whether to hire a male or female worker and the wage that they must pay means that conditional on gender, all workers are equally productive. The counterfactual dynamics I am interested in are at the level of occupation and gender, and productivity and utility are allowed to vary freely by occupation and gender. Therefore including individual worker productivity heterogeneity should not have a major impact on my results. In terms of biasing parameter estimates, I may overestimate how undesirable small occupations are. High wages in small occupations may be due to selecting the most skilled workers rather than dis-amenities. Similarly I may overestimate the value of amenities in large occupations if wages are low due to a lower than average level of worker skill, rather than attractive amenities.

@@

The payoff to the firm of not hiring any worker, $\pi^N_j$, is normalized to one, so that the log payoff to not filling a vacancy is zero, ($ \bar{\pi}^N_j = 0$).

@@

@@

Workers choose a job $j$ to maximize log utility $\bar{u}^{i}_j$.  The log utility parameters for working, $\bar{u}^g_o$, include the disutility of working at all and are expected to be negative. However the underlying utility parameters $u^g_o = exp(\bar{u}^g_o)$ will always be positive leading to complementarity in wage and non-wage utility. Similarly, ${\xi^g_j}$ will be assumed to be log-normally distributed and therefore always positive, so the higher the dis-amenity value of the job the lower the utility.

The common taste parameter for non-employment $\bar{u}^g_N$, where non-employment means either not in the labor force or unemployed, is normalized to zero. It is also assumed that in non-employment workers receive no wages or value from job amenities. The payoff to non-employment ($\bar{u}^i_N$) is therefore  $ \bar{\eta}^i_N $, the idiosyncratic taste for non-employment.

@@

So $\alpha^g_o$ and  $\gamma^g$ will be the parameters of interest, but $\bar{u}^{g}_o $ can be recovered from these parameters and the fraction female.

@@

@@

The fact that this wage vector supports the pairwise stable matching follows from the optimization problems of jobs and workers. Workers choose an occupation to maximize utility, and firms choose a worker to maximize rate of return, so the chosen job $j^*$ and worker $i^*$ respectively must satisfy

\section{Empirical Strategy} \label{empirical}
The empirical strategy consists of two stages: first I use maximum likelihood to disentangle the selection effects of worker and firm optimization on observed matches and wages. I am able to separately identify the willingness-to-pay of the firm and the reservation wage distributions of the workers. This effectively separates the two sides of the market, unlike previous applications of transferable utility matching models, such as \citeA{Choo2006} and \citeA{Chiappori2015}, who were only able identify total match surplus because they lacked data on transfers.\footnote{\citeA{Fox2008c} is able to estimate payoffs from the interaction of characteristics of both sides of the market but not full payoff functions.}

Second, the estimated reservation wage distributions from the maximum likelihood estimation are used to estimate the worker's utility. Workers care about non-wage amenities, wages, and the fraction female. In the second stage of estimation, a panel regression for each gender is run of shares of workers in each occupation on reservation wages, occupation intercepts, and fraction female. Instruments are used to control for omitted variables correlated with both fraction female and wages over time, and provide clean variation in the fraction female and reservation wage to trace out labor supply.

@@

For clarity of identification, I estimate the two sides of the market, workers and jobs, separately. The job side of the model is a maximum likelihood estimation exploiting the full variation of the observed wage distribution at the individual match level. The worker side, by contrast, is an instrumental variables regression at the occupation-year level, where instruments provide clean variation in the fraction female and reservation wage to trace out labor supply. Combining the two estimation steps using joint GMM is theoretically feasible and would have the advantage that all data moments are used to identify all parameters. However the large number of parameters (914 parameters at once)\footnote{There are 138 firm side parameters estimated in the first stage in each of 6 waves of data, for a total of 828 parameters. There are 86 worker parameters estimated in the second stage over all waves of data.} and the need to pool all years of data (over 7 million observations) makes joint estimation computationally unattractive.

@@

In the following sections I lay out the assumptions that allow the ML estimation of the willingness-to-pay and reservation wages. The reservation wages are then taken as data in the estimation of worker utility parameters, which is described in the following section.

@@

A firm will only hire a worker if the willingness-to-pay for that worker is higher than the wage the worker will accept in equilibrium, which means the payoff $ \pi^g_j $ must be greater than one.

There is no equivalent on the firm side to the idiosyncratic preference for non-employment that exists on the worker side. Heterogeneity in the outside option for a firm could reflect heterogeneity in the cost of hiring for different jobs, or the substitutability of a particular job in the firm production function. I make the assumption of a fixed outside option because then the model is identified without use of data on vacancies,\footnote{I can estimate the willingness-to-way of the firms with the maximum observed wage by gender and occupation. This strategy is less stable than using vacancy data.} allowing me to test robustness to the use of the vacancy data.

@@

\subsubsection{Likelihood Function}
%The parameters that we need to identify for the job side of the market are the type level market clearing common component of wages $W^g_o$, the willingness-to-pays $WTP^g_o$,  and the scale parameter of the job amenities shock $\xi^g_j$, which is $\sigma^g_{\xi}$.

Parameters of the model are

$$ \theta = \{ WTP^g_o,  \alpha^g_o, \gamma^g, \sigma^g_\eta, \sigma^g_\xi  \} $$

>>>>

\abstract{Women remain concentrated in certain occupations despite the drastic increase in women's labor force participation in the U.S. since 1960. I examine whether occupations remain segregated because workers prefer to enter occupations that already employ more of their own gender. I build a model of labor supply and demand in which firms care about the gender and wages of their employees, and workers get utility from their occupation, wage, and the number of women in their occupation. Using a Bartik instrumental variables strategy, I find that women prefer to enter into female-dominated occupations, but men show no evidence of gender preference. In a world without women's gender preference, occupations would be 50 percent less segregated and the gender wage gap 18 percent smaller. Homophily can mean that a given occupation could be either male or female dominated depending on how many women were in it initially. However, women's preference for working with women would have to be almost twice as strong for the number of women in the past to affect whether an occupation is male or female dominated today.}

@@

@@

@@

@@

%% put firm STUFF HERE

%unlike the marriage market, it is not clear what it means for a job to be ``unmatched". Theoretically, the number of unmatched jobs would be the number of jobs that would exist if labor were free. Rather than estimate production functions and capacity constraints by occupation, I choose to use data on posted vacancies that remain unfilled for some time.

%The decision to post a vacancy is endogenous to market conditions, so vacancy data is an imperfect proxy for unfilled jobs.

%In order to limit the importance of the vacancy data in my estimation, I choose a specification that in principle does not require it, by assuming non-negative profit on the job side. However in practice this relies very heavily on the functional form assumption on the tail of the distribution of job dis-amenities. As a result in most specifications I use data on vacancies by occupation imputed from the Job Openings and Labor Turnover Survey \cite{JOLTS}.

@@

@@

@@

The payoff to the firm of not hiring any worker, $\pi^N_j$, is normalized to one, so that the log payoff to not filling a vacancy is zero, ($ \bar{\pi}^N_j = 0$). There is no equivalent on the firm side to the idiosyncratic preference for non-employment that exists on the worker side. Heterogeneity in the outside option for a firm could reflect heterogeneity in the cost of hiring for different jobs, or the substitutability of a particular job in the firm production function.

I make the assumption of a fixed outside option because then the model is identified without use of data on vacancies, allowing me to test robustness to the use of the vacancy data. I can estimate the willingness-to-way of the firms with the maximum observed wage by gender and occupation. This strategy is less stable than using vacancy data. A firm will only hire a worker if the willingness-to-pay for that worker is higher than the wage the worker will accept in equilibrium, which means the payoff $ \pi^g_j $ must be greater than one.

The fact that jobs care only about whether to hire a male or female worker and the wage that they must pay means that conditional on gender, all workers are equally productive. This assumption is important because it allows me to separate the matching problem into two separate discrete choice problems, one for each side of the market \cite{Galichon2013}. The counterfactual dynamics I am interested in are at the level of occupation and gender, and productivity and utility are allowed to vary freely by occupation and gender. Therefore including individual worker productivity heterogeneity should not have a major impact on my results. In terms of biasing parameter estimates, I may overestimate how undesirable small occupations are. High wages in small occupations may be due to selecting the most skilled workers rather than dis-amenities. Similarly I may overestimate the value of amenities in large occupations if wages are low due to a lower than average level of worker skill, rather than attractive amenities.
% JUSTIFY MORE WHY ORTHOGONAL

@@

Workers choose an occupation at the beginning of their lives, based on their individual tastes for occupations and the wage offer for their gender in that occupation. The job choice is binding for life.\footnote{Since there is no switching jobs, I assume that education choice is tied to job choice and do not treat education as a separate decision.} I assume that workers are myopic and do not anticipate future changes in the market that would lead them to want to switch occupations.\footnote{Lowering switching cost would lead to faster convergence to a fixed point in the fraction female by occupation. However, relaxing myopia would pose a challenge to tractability because the endogenous attribute (fraction female) would then depend on expectations over the future behavior of all agents, complicating today's choice set considerably.}

@@

Workers choose a job $j$ to maximize log utility $\bar{u}^{i}_j$. The common taste parameter for non-employment $\bar{u}^g_N$, where non-employment means either not in the labor force or unemployed, is normalized to zero. It is also assumed that in non-employment workers receive no wages or value from job amenities. The payoff to non-employment ($\bar{u}^i_N$) is therefore  $ \bar{\eta}^i_N $, the idiosyncratic taste for non-employment.\footnote{The log utility parameters for working, $\bar{u}^g_o$, include the disutility of working at all and are expected to be negative. However the underlying utility parameters $u^g_o = exp(\bar{u}^g_o)$ will always be positive leading to complementarity in wage and non-wage utility. Similarly, ${\xi^g_j}$ will be assumed to be log-normally distributed and therefore always positive, so the higher the dis-amenity value of the job the lower the utility. }

@@

So $\alpha^g_o$ and  $\gamma^g$ will be the parameters of interest, but $\bar{u}^{g}_o $ can be recovered from these parameters and the fraction female. %% PUT IN FOOTNOTE why sigmas matter, but CUT THIS
Lastly, I estimate four variance parameters in total: variance of taste for occupation for men and women, and variance of job dis-amenities for men and women. These parameters are important because they govern the elasticity of labor supply to changes in wages or other non-wage amenities such as the gender ratio in the counterfactual, as well as the demand response to changes in equilibrium wage. Unlike \citeA{Chiappori2015}, who uses many markets to identify these parameters, I use the observed wage distribution as suggested in \citeA{Salanie2014a}. However I estimate using maximum likelihood in order to use the information from the wage distribution efficiently, and because the lognormal assumption means that the scale parameters cannot be separately estimated from all other parameters, as noted above. For all these reasons maximum likelihood is the most natural strategy for my application.

@@

% REPLACE
The lifetime income distributions by gender and occupation needed to clear the market are determined in equilibrium to equate demand and supply, and are made up of two components: a job-specific component, and an aggregate component that varies by occupation and gender. The job-specific component reflects the idiosyncratic disutility of taking a specific job. In equilibrium firms must make workers indifferent across individual jobs within an occupation, and therefore firms perfectly compensate workers for idiosyncratic disutility. Wage heterogeneity within occupation and gender emerges from these job-specific compensating differentials.

@@

The fact that this wage vector supports the pairwise stable matching follows from the optimization problems of jobs and workers. See appendix \ref{sec.Stability} for proof of stability.

\section{Empirical Strategy} \label{empirical}

Previous literature only identifies the sum of wage and non-wage utility. Because a key model parameter is how non-wage utility varies with occupation fraction female, it is critical that I separately identify non-wage utility from utility from wages. The identification strategy in this paper builds off empirical applications of matching models to marriage markets (e.g. \citeA{Choo2006} and \citeA{Chiappori2015}), but differs in several key respects.

% REPLACE
%The difficulty is that observetd transfers are the sum of an idiosyncratic and an aggregate component, as discussed above, and for counterfactuals these must be separately identified. The observed distribution of wages is also the result of optimization over these components and other model parameters to be estimated, specifically the willingness-to-pay for workers by gender and the variance of the distribution of unobserved heterogeneity in job amenities.

The first step towards separate identification is to assume that observed wages are the transfers between job and worker that clear the market. Although some transfers in the labor market may be non-wage, such as enhanced benefits, wages are a natural first-order approximation. Previous literature applying transferable utility to the marriage market has assumed a logit assumption for unobserved transfers because logit is most tractable. I model wages as lognormal in order to match the observed wage distribution, which is greater than zero and has a long right tail.\footnote{Unlike the logit case, the variance of the distribution of the maximum of lognormals depends on other model parameters to be estimated, so the variance needs to be jointly estimated unlike in \citeA{Salanie2014a} for example.} Second, I define an unmatched worker as a workers who is unemployed or out of the labor force, and an unmatched job as a posted vacancy that has remained unfilled for a period of time. Since vacancy data is potentially a poor proxy for what it means for a job to be unmatched, I try to limit the importance of the vacancy data in my estimation procedure by choosing a specification that in principle does not require it, by assuming non-negative profit on the job side.

%The empirical strategy consists of two stages: first I use maximum likelihood to disentangle the selection effects of worker and firm optimization on observed matches and wages. I am able to separately identify the willingness-to-pay of the firm and the reservation wage distributions of the workers. This effectively separates the two sides of the market, unlike previous applications of transferable utility matching models, such as \citeA{Choo2006} and \citeA{Chiappori2015}, who were only able identify total match surplus because they lacked data on transfers.\footnote{\citeA{Fox2008c} is able to estimate payoffs from the interaction of characteristics of both sides of the market but not full payoff functions.}

%Second, the estimated reservation wage distributions from the maximum likelihood estimation are used to estimate the worker's utility. Workers care about non-wage amenities, wages, and the fraction female. In the second stage of estimation, a panel regression for each gender is run of shares of workers in each occupation on reservation wages, occupation intercepts, and fraction female. Instruments are used to control for omitted variables correlated with both fraction female and wages over time, and provide clean variation in the fraction female and reservation wage to trace out labor supply.

@@

For clarity of identification, I estimate the two sides of the market, workers and jobs, separately. The job side of the model is a maximum likelihood estimation exploiting the full variation of the observed wage distribution at the individual match level. The worker side, by contrast, is an instrumental variables regression at the occupation-year level, where instruments provide clean variation in the fraction female and reservation wage to trace out labor supply.\footnote{Combining the two estimation steps using joint GMM is theoretically feasible and would have the advantage that all data moments are used to identify all parameters. However the large number of parameters (914 parameters at once) and the need to pool all years of data (over 7 million observations) makes joint estimation computationally unattractive.}

% \footnote{There are 138 firm side parameters estimated in the first stage in each of 6 waves of data, for a total of 828 parameters. There are 86 worker parameters estimated in the second stage over all waves of data.}

@@

%%%%%In the following sections I lay out the assumptions that allow the ML estimation of the willingness-to-pay and reservation wages. The reservation wages are then taken as data in the estimation of worker utility parameters, which is described in the following section.

@@

@@

%\subsubsection{Likelihood Function}
%%The parameters that we need to identify for the job side of the market are the type level market clearing common component of wages $W^g_o$, the willingness-to-pays $WTP^g_o$,  and the scale parameter of the job amenities shock $\xi^g_j$, which is $\sigma^g_{\xi}$.
%
%Parameters of the model are
%
%$$ \theta = \{ WTP^g_o,  \alpha^g_o, \gamma^g, \sigma^g_\eta, \sigma^g_\xi  \} $$
%
%Recall that $WTP^g_o$ are the firms' willingness-to-pay parameters, $\alpha^g_o$ the workers' non-wage utility unrelated to the fraction female, $\gamma^g$ the value of the fraction female to the worker, $\sigma^g_\eta$ the scale of the worker taste heterogeneity, and $\sigma^g_\xi$ the scale of the job dis-amenities heterogeneity. The unconditional centers of the reservation wage distributions, $W^g_o$, are reduced form outcomes determined in equilibrium as a function of the primitives $\theta$. Nevertheless recovering the $W^g_o$ from the observed conditional wage distributions is critical to recovering the fundamental parameters in the worker's utility function.
%

%%%Recall the log wage specification, $$ \widebar{Wage}^g_j = \bar{W}^g_o -  \bar{\xi}^g_j $$
